Numerical Analysis \\ Assignment 9

### 2017-10-28 Graduate - Numerical Analysis/Assignment 10/assignment.tex
add prob 3, 4

\problem{3}{Problem 4.21, Page 243}

\problem{4}{Problem 4.23, Page 243}

\end{document}

>>>>

\problem{3}{Problem 4.21, Page 243}
\solution{Proof}{
Denote $\varphi_n(x) = A_n x^n+B_n x^{n-1}+\cdots $, and $A_n > 0$. Then by (4.4.21), 
$$
\varphi_{n+1}(x) = (a_n x+b_n)\varphi_n(x)-c_n\varphi_{n-1}(x).
$$
We add a $\varphi_0(x)$ to this series, and $\varphi_{0}(x) = A_0 > 0$.
First, when $n = 1$, since
$$
\int_{a}^{b} w(x)\varphi_1(x)\varphi_0(x)dx = \int_{a}^{b}A_0w(x)\varphi_1(x)dx = 0 = A_0\varphi_1(\xi)\int_{a}^{b}w(x) dx,
$$
we know $\varphi_1(\xi) = 0, ~\xi\in (a, b)$. Then we show $\varphi_2(x)$ has two different roots in $(a, b)$. First,
$$
\varphi_2(\xi) = (a_1\xi+b_1)\varphi_1(\xi)-c_1\varphi_0(\xi) = -c_1\varphi_0(\xi) < 0.
$$
Suppose $\varphi_2(x)$ does not change sign in $(a, b)$, then
$$
\int w(x)\varphi_2(x)\varphi_0(x) = A_0\int w(x) \varphi_2(x) < 0.
$$
It makes a contradiction with orthogonality. Then there $\exists x_1\in(a, b) $, s.t. $\varphi_2(x_1) = 0$. Since $A_2 > 0$, $x_1$ cannot be a double root. If $\varphi_2(x)$ has only one root in $(a, b)$, then 
$$
\varphi_2(x)(x-x_1) = q(x)(x-x_1)^2,
$$
integrate by $w(x)$, then since $(x-x_1)$ is of degree 1, left side is 0. But we know $q(x)$ has no root in $(a, b)$, it does not change sign in $(a, b)$, thus the integration is not 0. It makes a contradiction. Thus $\varphi_2(x)$ has two different roots in $(a, b)$, and with $\varphi_2(\xi) < 0$ and $A_2 > 0$, we know the two roots are in $(a, \xi)$ and $(\xi, b)$ seperately. Using the same method, we know that $\varphi_n(x)$ has $n$ different roots in $(a, b)$.

Now we assume this proposition holds for $\varphi_m(x)$, $m \le n$. Denote roots of $\varphi_{n}(x)$ to be $x_i$, then since
$$
\varphi_{n+1}(x_i) = (a_n x_i+b_n)\varphi_n(x_i)-c_n\varphi_{n-1}(x_i) = -c_n\varphi_{n-1}(x_i),
$$
$$
\varphi_{n+1}(x_{i+1}) = (a_n x_{i+1}+b_n)\varphi_n(x_{i+1})-c_n\varphi_{n-1}(x_{i+1}) = -c_n\varphi_{n-1}(x_{i+1}),
$$
from the assumption we know $\varphi_{n-1}(x_i) $ and $\varphi_{n-1}(x_{i+1}) $ has different signs, which means $\varphi_n(x) $ has a root in each of this intervals. Then from $\varphi_n $ has $n$ different roots in $(a, b)$, from induction we get the result.
}

\problem{4}{Problem 4.23, Page 243}
\solution{Solution}{
We know the orthogonal functions on $[-1, 1]$ with weight function $w(x) = \frac{1}{\sqrt{1-x^2}}$ are Chebyshev polynomials:
$$
T_0(x) = 1, ~T_1(x) = x, ~T_2(x) = 2x^2-1.
$$
For any $p(x) = a_0+a_1x+a_2x^2 $ of degree 2 that minimizes the distance, we have
$$
a_j = (f, T_j), ~j = 0, 1, 2.
$$
Thus
$$
a_0 = (f, T_0) = \int_{-1}^{1} \frac{1}{\sqrt{1-x^2}}\cos^{-1}xdx = \int_{\pi}^{0}-ydy = \frac{\pi^2}{2},
$$
$$
a_1 = (f, T_1) = \int_{-1}^{1}\frac{x}{\sqrt{1-x^2}}\cos^{-1}xdx = \left.\int_{0}^{\pi} y\cos ydy = y\sin y\right|_{0}^{\pi} - \int_{0}^{\pi} \sin ydy = -2,
$$
$$
a_2 = (f, T_2) = \int_{-1}^{1}\frac{2x^2-1}{\sqrt{1-x^2}}\cos^{-1}xdx = \left.\int_{0}^{\pi}y\cos 2ydy = \frac{y\sin 2y}{2}\right|_{0}^{\pi} - \frac{1}{2}\int_{0}^{\pi}\sin 2ydy = -\frac{1}{2}.
$$
Thus
$$
p_2(x) = -\frac{1}{2}x^2-2x+\frac{\pi^2}{2}.
$$
}
\end{document}
